---
title: "iris"
output: github_document
Author: "Bhasheyam Krishnan"
---
Loading the dataset
```{r}
irirdata = read.csv(file = "iris.csv")
```

```{r}
fix(irirdata)
```

check for missing data
```{r}
irirdata[irirdata == ""] = NA
sum(is.na(irirdata))
```

The dataset deosnt have any missing data.
```{r}
summary(irirdata)
```

From the above we can see all nuemeric type and one is of the type class. So the it is the type of classification


so lets scale the the data .

```{r}
library(AppliedPredictiveModeling)
transparentTheme(trans = .4)
library(caret)
featurePlot(x = irirdata[, 2:5], 
            y = irirdata$Species, 
            plot = "pairs",
            ## Add a key at the top
            auto.key = list(columns = 3))
```
```{r}
library(AppliedPredictiveModeling)
library(caret)
transparentTheme(trans = .9)
featurePlot(x = iris[, 1:4], 
            y = iris$Species,
            plot = "density", 
            ## Pass in options to xyplot() to 
            ## make it prettier
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")), 
            adjust = 1.5, 
            pch = "|", 
            layout = c(4, 1), 
            auto.key = list(columns = 3))
```
```{r}
library(AppliedPredictiveModeling)
library(caret)
featurePlot(x = irirdata[, 2:5], 
            y = irirdata$Species, 
            plot = "box", 
            ## Pass in options to bwplot() 
            scales = list(y = list(relation="free"),
                          x = list(rot = 90)),  
            layout = c(4,1 ), 
            auto.key = list(columns = 2))
```


except sepalWidth feature all other have seperate cluster.

```{r}
normal = function(x){
  num = x - min(x)
  den =max(x) - min(x)
  return ( num/den)
  
}
iris_scaled = as.data.frame(lapply(Filter(is.numeric,irirdata),normal))
iris_scaled = cbind(iris_scaled,Species =irirdata[,6])
summary(iris_scaled)
```


Now data is prepared and lets look for outliers.
```{r}
boxplot(iris_scaled[,2:5])


```
Only Sepalwidth has the outliers 4 of them are there.
```{r}
summary(iris_scaled)
```

```{r}
iris_scaled = iris_scaled[iris_scaled$SepalWidthCm < 0.80,]
```

```{r}
dim(iris_scaled)
```
outliers are removed

Spliting the data into train and test.

```{r}
set.seed(123)
jumble = runif(nrow(iris_scaled))
iris_scaled = iris_scaled[ordered(jumble),]
sampleindex = sample(2,nrow(iris_scaled),replace = TRUE,prob = c(0.67,0.33))
iris_train = iris_scaled[sampleindex == 1,]
iris_test = iris_scaled[sampleindex == 2,]
dim(iris_test)
dim(iris_train)
```


```{r}
library(caret)
library(rpart.plot)


trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

dtree_fit <- train(Species ~., data = iris_train[,2:6], method = "rpart",
                   parms = list(split = "information"),
                   trControl=trctrl,
                   tuneLength = 10)



```


```{r}
library(rpart.plot)
prp(dtree_fit$finalModel, box.palette = "Reds", tweak = 1.2)
```
```{r}
decisionpred = predict(dtree_fit,iris_test[,2:5])
confusionMatrix(iris_test[,6],decisionpred)
```
```{r}
iris_train[,6] = as.factor(iris_train[,6])
iris_test[,6] = as.factor(iris_test[,6])
```


KNN model 
```{r}
library(caret)
cctrl <- trainControl(method = "cv", number = 3, repeats = 3)

knnmodel <- train(Species ~ ., data = iris_train[,2:6], 
                            method = "knn", 
                            trControl = cctrl,
                             tuneLength = 15
                            )

```


```{r}
knnpred = predict(knnmodel, iris_test[,2:6])
confusionMatrix(knnpred,iris_test[,6])
```

```{r}
names(iris_test)
```
#```{r}
#####
library(plyr)

plotdata = cbind(x = iris_test$SepalLengthCm, y = iris_test$PetalLengthCm , predicted = knnpred)

find_hull = function(df) df[chull(df$x, df$y), ]

boundary = ddply(plotdata, .variables = "predicted", .fun = find_hull)

#```


```{r}
library(ggplot2)
ggplot(iris_test, aes(x=SepalLengthCm,y =PetalWidthCm,color = knnpred))+ geom_point(size = 3) + ggtitle("knnmodel") 

```

```{r}
library(AppliedPredictiveModeling)
transparentTheme(trans = .4)
library(caret)
featurePlot(x = iris_test[, 2:5], 
            y = knnpred, 
            plot = "pairs",
            ## Add a key at the top
            auto.key = list(columns = 3))
```

The above is the visiualization of test results